### 2024-11-15 paper/paper.qmd
add paper

---
title: "Sale difference between Loblaws and Voila"
author: 
  - Viet Nguyen
  - Doorwan Wang
  - Yihang Xu
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "Sales promotions are widely used by grocery vendors to attract customers with promises of discounts and savings. However, there is an increasing concern that some of these promotions may not represent genuine price reductions but instead involve pre-sale price inflation—an artificial increase in price just before a sale, creating an illusion of a discount when the price returns to its original level. This research investigates the prevalence of such practices by examining historical price data across a range of grocery products and vendors. By analyzing price trends leading up to and following sales events, we aim to identify patterns of potential price manipulation. Specifically, we assess whether prices exhibit significant increases just prior to advertised sales and how post-sale prices compare to both pre-sale and long-term average prices. Our findings will offer insights into the authenticity of grocery sale discounts, empowering consumers with knowledge to make informed purchasing decisions and encouraging retailers toward transparent pricing practices. This study contributes to a broader understanding of pricing strategies in retail and the implications of promotional tactics on consumer trust."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)

data <- read_csv("../data/clean_data.csv")

```

# Introduction
Sales promotions have long been a cornerstone of retail marketing, with grocers frequently offering "discounts" that claim to provide customers with significant savings. However, some consumers and industry analysts have raised questions about the authenticity of these discounts, speculating that prices may be artificially inflated shortly before a sale, only to be reduced back to their original or slightly lower levels during the promotion. This practice, often referred to as "price jacking," can give the illusion of a significant discount while actually offering little to no real savings.

This research explores the question: **When something's on "sale," was the price artificially increased just before the sale, only to be lowered back to its regular level?** To investigate this, we will analyze historical price data from various grocery vendors to examine patterns preceding and following sales events. By comparing regular and sale prices over time, this study aims to identify any indications of pre-sale price increases, offering insight into whether certain sales are truly beneficial for consumers or merely marketing tactics designed to create a perception of value.

The findings from this research will help consumers better understand the nature of grocery discounts and offer guidance on identifying genuine sales. Additionally, the results will provide insights for retailers into the long-term trust implications of their pricing strategies and potential opportunities for increased transparency.

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

```{r}
#| echo: false
#| warning: false
#| message: false
print(data)
```

```{r}
#| echo: false
#| warning: false
#| message: false
data_long <- data %>%
  mutate(price_type = ifelse(!is.na(other), "Sale", "Regular")) %>%
  pivot_longer(cols = c(current_price, old_price), values_to = "price")

# Create the line plot
ggplot(data_long, aes(x = product_vendor, y = price, color = price_type, linetype = price_type, group = interaction(price_type, product_vendor))) +
  geom_line() +
  geom_point() +
  facet_wrap(~product_vendor) + # Separate lines for each vendor
  labs(title = "Price Comparison Over Time for Apples Ambrosia",
       x = "Vendor", 
       y = "Price", 
       color = "Price Type", 
       linetype = "Price Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Measurement
To investigate the potential for artificial price increases prior to sales events, this study employs a data-driven measurement approach to track and analyze price fluctuations over time. The measurement strategy is broken down as follows:

1. **Baseline Price Identification**: For each product in the dataset, we identify the "baseline price" as the most frequently observed price outside of any sale periods. This price will serve as a reference point for measuring any deviations leading up to and following sale events.

2. **Sale Periods and Regular Periods**: We classify each time period as either a "sale period" or "regular period" based on the presence of a discount. Sale periods are defined as those where a price reduction is advertised, with "old price" values recorded in the dataset to indicate a discount.

3. **Price Inflation Detection**: To determine if price increases precede sale events, we measure the average price in the 30 days leading up to each sale period and compare it to the baseline price. We record any instances where pre-sale prices exceed the baseline by a significant margin (e.g., more than 5%).

4. **Post-Sale Price Comparison**: After the sale event, we compare the product’s price to both the baseline and pre-sale prices to determine if the price returns to a normal level, remains higher, or is permanently reduced. This comparison helps assess whether the sale was a genuine discount or if it merely restored the price to a regular level.

5. **Statistical Analysis of Patterns**: We use statistical tests, such as paired t-tests or Wilcoxon signed-rank tests, to analyze whether observed price increases before sales are significant across vendors and product families. Additionally, we track the frequency and magnitude of pre-sale price increases across products and vendors to identify any patterns.

Through these measurements, we aim to rigorously test the hypothesis that prices are artificially inflated prior to sales, providing a comprehensive view of price dynamics around promotional periods.

# Results
Price Analysis of "Apples Ambrosia" Across Vendors
The dataset consists of price data for Apples Ambrosia from two grocery vendors: Voila and Loblaws. For each vendor, prices are observed during periods labeled as "SALE" (Voila) and regular pricing periods (Loblaws). The analysis investigates the following:

Price Comparison Between Vendors:

Voila offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items. This suggests a $1.00 decrease in price during the sale period, indicating that Voila is promoting a sale by reducing the price from the regular price.
Loblaws, on the other hand, has a consistent current price of $5.49 for all data points. This is lower than the sale price at Voila. Additionally, Loblaws’ old price is slightly higher, averaging around $5.49 to $6.58. However, the regular price at Loblaws appears to be stable without the promotional fluctuations observed at Voila.
Sale Behavior Analysis at Voila:

At Voila, the old price for the apples is $6.49, which is advertised as the regular price. The current price during the sale is lower by exactly $1.00, consistently across multiple observations. This consistent pricing pattern across the dataset suggests that Voila uses a fixed price reduction strategy during sales, offering a price decrease of $1.00 off the original price during promotions.
An important observation is that the old price at Voila (before the sale) is consistently $6.49, which may suggest that prices were artificially elevated just ahead of the sale, only to be reduced back down to the normal sale price during the promotional period. The stable difference between the sale and non-sale prices raises a question about the pricing practices at Voila.
Price Consistency at Loblaws:

At Loblaws, no specific sale event is recorded, and the prices are relatively stable, with current prices consistently at $5.49 for "Apples Ambrosia." The old price for Loblaws varies slightly, but the regular price does not undergo significant fluctuations like the sales at Voila. This stability in pricing indicates that Loblaws may not engage in price manipulation strategies for sale periods as seen at Voila. The current price of $5.49 is relatively consistent and does not appear to be influenced by any temporary sale-like conditions.
Conclusion on Price Manipulation:

Based on the price data, Voila may be engaging in price manipulation practices, where the price of the apples is first elevated to $6.49 and then temporarily reduced to $5.49 during the sale period, creating an illusion of a larger discount. This could be viewed as an artificial price increase before a sale, which aligns with the research question regarding whether prices are jacked up before a sale and lowered back to normal.
Loblaws, by contrast, maintains a more stable pricing structure and does not appear to engage in such pricing practices.
Overall, this analysis suggests that Voila may be artificially inflating prices ahead of sales to make the sale price appear more significant, while Loblaws maintains a consistent price strategy without such fluctuations. Further statistical analysis, such as testing for significant price changes before and after sales, would be necessary to substantiate this hypothesis more robustly.

# Discussion
The analysis of price data for **Apples Ambrosia** from **Voila** and **Loblaws** provides insight into pricing strategies across two grocery vendors. Specifically, the research question—whether prices are artificially increased just before a sale, only to be lowered back down—was examined by comparing **current prices** and **old prices** during sale periods at Voila and regular pricing at Loblaws.

## Price Manipulation at Voila

The results suggest that **Voila** may be engaging in a pricing strategy that aligns with the hypothesis of artificially inflating prices ahead of a sale. At Voila, the **old price** for the apples is consistently **$6.49**, which is advertised as the regular price. However, the **current price** during the sale period is consistently lower at **$5.49**, resulting in a fixed **$1.00 discount** across all sale entries. This consistent reduction implies that **Voila's sales** may be structured to create the illusion of a larger price cut, as prices are raised to **$6.49** just before the sale and then reduced to the normal price of **$5.49** during the sale.

This pricing behavior may lead customers to believe they are receiving a more significant discount than they actually are. Such price adjustments raise questions about transparency in pricing and the ethical implications of using these sales strategies. The strategy of artificially inflating prices before a sale is a tactic often criticized in retail, as it may mislead consumers into thinking they are getting better deals than they truly are. While the **$1.00 discount** in the sale period does represent a legitimate price reduction, it may not be as substantial as it appears when compared to the price manipulation in the lead-up to the sale.

## Consistent Pricing at Loblaws

In contrast, **Loblaws** maintains a more consistent pricing structure. The **current price** of **$5.49** for **Apples Ambrosia** is stable across all observations, and there is no recorded **sale** price or fluctuation in price over time. The **old price** at Loblaws varies slightly, but the difference between the **current** and **old price** is not as pronounced as at Voila. This suggests that Loblaws does not engage in the same type of artificial price elevation seen at Voila. Their pricing appears more transparent, without the pre-sale price hikes that would create the illusion of larger discounts.

The stable pricing at **Loblaws** could be interpreted as a more consumer-friendly approach, with less likelihood of misleading customers about the "value" of sales. In fact, this could suggest that **Loblaws** may not use sales as a primary strategy to attract customers, but instead maintains competitive and consistent pricing for everyday purchases. While the price at Loblaws is lower than the sale price at Voila, it is important to consider that Loblaws may be applying a different pricing strategy based on its market positioning, competitive environment, and customer base.

## Implications for Consumers

From a consumer perspective, the price manipulation strategy observed at Voila may have broader implications. Customers who are aware of the practice may become skeptical of sales and discounts, leading them to question the legitimacy of promotional offers. Such skepticism could erode trust in the vendor and result in a loss of customer loyalty. On the other hand, the consistent pricing at Loblaws could build consumer trust, as shoppers may feel confident that the price they see is the price they pay, without the manipulation of artificial sales tactics.

However, it is also worth noting that the **$1.00 price difference** observed between **sale** and **regular price** is relatively modest and might not drastically affect consumer behavior unless similar strategies are applied to a wide range of products. Thus, while the price manipulation issue is noteworthy, it may not be the sole factor driving purchasing decisions in the broader context of grocery shopping, where convenience, availability, and other factors also play a significant role.

## Limitations and Further Research

While this study provides useful insights, there are limitations to the analysis. The dataset only includes a limited range of data points for **Apples Ambrosia** across two vendors, which may not fully represent the broader pricing strategies employed by the vendors for all product categories. Further research would benefit from a more comprehensive dataset, encompassing a variety of products, seasons, and vendors. This would allow for a more robust understanding of pricing practices across different contexts.

Moreover, additional statistical analysis could be conducted to test whether the observed price differences are statistically significant. This would help quantify the extent to which prices are artificially inflated before sales and provide more concrete evidence for or against the practice.

## Conclusion

In conclusion, the analysis indicates that **Voila** may be engaging in artificial price elevation ahead of sales, creating an illusion of a larger discount when the price is lowered back to normal during the sale period. **Loblaws**, in contrast, maintains a consistent and stable pricing strategy. While this raises important questions about transparency and fairness in pricing practices, it also provides an opportunity for consumers to make more informed decisions by understanding how sale prices may be manipulated. Further research would be needed to investigate this practice across a broader range of products and vendors, with an emphasis on consumer perceptions and market dynamics.

# Appendix {-}

# References

### 2024-11-15 paper/paper.qmd
update paper.qmd

@@
 data <- read_csv("../data/clean_data.csv")
-
 ```
@@
 
-This research explores the question: **When something's on "sale," was the price artificially increased just before the sale, only to be lowered back to its regular level?** To investigate this, we will analyze historical price data from various grocery vendors to examine patterns preceding and following sales events.
+This research explores the question: When something's on "sale," was the price artificially increased just before the sale, only to be lowered back to its regular level?** To investigate this, we will analyze historical price data from various grocery vendors to examine patterns preceding and following sales events.
 By comparing regular and sale prices over time, this study aims to identify any indications of pre-sale price increases, offering insight into whether certain sales are truly beneficial for consumers or merely marketing tactics designed to create a perception of value.
@@
 
-# Appendix {-}
-
 # References

### 2024-11-15 paper/paper.qmd
update files

@@
 library(tidyverse)
+library(readr)
 library(ggplot2)
@@
 library(tibble)
+library(knitr)
 
@@
 ## Overview
-
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-print(data)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-data_long <- data %>%
-  mutate(price_type = ifelse(!is.na(other), "Sale", "Regular")) %>%
-  pivot_longer(cols = c(current_price, old_price), values_to = "price")
-
-# Create the line plot
-ggplot(data_long, aes(x = product_vendor, y = price, color = price_type, linetype = price_type, group = interaction(price_type, product_vendor))) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(~product_vendor) + # Separate lines for each vendor
-  labs(title = "Price Comparison Over Time for Apples Ambrosia",
-       x = "Vendor", 
-       y = "Price", 
-       color = "Price Type", 
-       linetype = "Price Type") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
+We use the statistical programming language R [@citeR]....
+The data used in this paper came from the [@hammer].
+Following @tellingstories, we considerto capture the pricing dynamics of Ambrosia apples across different vendors, we extracted relevant data from a larger dataset containing multiple products and vendors.
+The initial dataset included information from seven vendors, each offering various products.
+Our focus was on the specific product, "Ambrosia Apples," which was sold by two vendors—NoFrills and Loblaws.From this comprehensive dataset, we filtered out all other products, isolating only the entries related to Ambrosia apples.
+The extracted data included key information such as the product ID, product name, current price, and old price.
+Additionally, a field labeled "other" contained further notes on sale conditions or discounts.
+The dataset also recorded timestamps, allowing us to track pricing changes over time.
 
 ## Measurement
+
 To investigate the potential for artificial price increases prior to sales events, this study employs a data-driven measurement approach to track and analyze price fluctuations over time.
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-loblaws_data
+#| tbl-cap: "Ambrosia Apples Data of Loblaws"
+#| warning: false
+
+data |>
+  filter(product_vendor == "Loblaws") |>
+  select(product_name,product_vendor,current_price,old_price,nowtime) |>      # Select the first 3 columns
+  kable(
+    col.names = c("Product Name","Product Vendor","Product Price(current)","Product Price(old)","Date"),
+    booktabs = TRUE
+  )
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-nofrills_data
+#| tbl-cap: "Ambrosia Apples Data of NoFrills"
+#| tbl-width: 2
+#| tbl-height: 2
+#| warning: false
+
+data |>
+  filter(product_vendor == "NoFrills") |>
+  select(product_name,product_vendor,current_price,old_price,nowtime) |>    
+  kable(
+    col.names = c("Product Name","Product Vendor","Product Price (current)","Product Price (old)","Date"),
+    booktabs = TRUE
+  )
+```
+
+The transition from real-world phenomena (e.g., pricing decisions and promotional strategies) to dataset entries was facilitated by these fields.Current and old prices reflected the vendors' listed prices at the time of data collection.
+Notes in the "other" field provided contextual information about whether the price was marked as a sale or if there were discounts.
+The "nowtime" column captured the exact timing of data collection, enabling us to analyze price fluctuations and assess sale validity over time.
+
+By focusing on two vendors and this specific product, we created a clean, targeted dataset shows in @tbl-loblaws_data and @tbl-nofrills_data that accurately reflects the pricing behavior for Ambrosia apples at NoFrills and Loblaws, laying the foundation for further analysis.
+
+## Outcome variables
+
+```{r}
+#| label: fig-compare_price_vendor
+#| fig-cap: Current and Old Prices of Ambrosia Apples Over Time
+#| echo: false
+#| warning: false
+#| message: false
+
+data$nowtime <- as.POSIXct(data$nowtime, format="%Y-%m-%d %H:%M:%S")
+
+data$product_vendor <- factor(data$product_vendor, levels = c("NoFrills", "Loblaws"))
+
+# Create the line plot
+ggplot(data, aes(x = nowtime)) +
+  geom_line(aes(y = current_price, color = product_vendor, linetype = "Current Price")) +
+  geom_line(aes(y = old_price, color = product_vendor, linetype = "Old Price")) +
+  labs(title = "Current and Old Prices of Ambrosia Apples Over Time",
+       x = "Date",
+       y = "Price ($)",
+       color = "Vendor",
+       linetype = "Price Type") +
+  theme_minimal() +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme(legend.position = "right")
+```
+We compared the apple prices of nofrills and loblaws at different times (including the comparison of old prices, where NoFrills does not include any old prices because this vendor does not have any discounts for this product and always maintains the original price), which can be seen in @fig-compare_price_vendor.
+
 # Results
-Price Analysis of "Apples Ambrosia" Across Vendors
-The dataset consists of price data for Apples Ambrosia from two grocery vendors: Voila and Loblaws.
-For each vendor, prices are observed during periods labeled as "SALE" (Voila) and regular pricing periods (Loblaws).
+
+## Price Analysis of “Apples Ambrosia” Across Vendors
+
+The dataset consists of price data for Apples Ambrosia from two grocery vendors: NoFrills and Loblaws.
+For each vendor, prices are observed during periods labeled as “SALE” (NoFrills) and regular pricing periods (Loblaws).
 The analysis investigates the following:
 
-Price Comparison Between Vendors:
-
-Voila offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items.
-This suggests a $1.00 decrease in price during the sale period, indicating that Voila is promoting a sale by reducing the price from the regular price.
-Loblaws, on the other hand, has a consistent current price of $5.49 for all data points.
-This is lower than the sale price at Voila.
-Additionally, Loblaws’ old price is slightly higher, averaging around $5.49 to $6.58.
-However, the regular price at Loblaws appears to be stable without the promotional fluctuations observed at Voila.
-Sale Behavior Analysis at Voila:
-
-At Voila, the old price for the apples is $6.49, which is advertised as the regular price.
-The current price during the sale is lower by exactly $1.00, consistently across multiple observations.
-This consistent pricing pattern across the dataset suggests that Voila uses a fixed price reduction strategy during sales, offering a price decrease of $1.00 off the original price during promotions.
-An important observation is that the old price at Voila (before the sale) is consistently $6.49, which may suggest that prices were artificially elevated just ahead of the sale, only to be reduced back down to the normal sale price during the promotional period.
-The stable difference between the sale and non-sale prices raises a question about the pricing practices at Voila.
-Price Consistency at Loblaws:
-
-At Loblaws, no specific sale event is recorded, and the prices are relatively stable, with current prices consistently at $5.49 for "Apples Ambrosia." The old price for Loblaws varies slightly, but the regular price does not undergo significant fluctuations like the sales at Voila.
-This stability in pricing indicates that Loblaws may not engage in price manipulation strategies for sale periods as seen at Voila.
-The current price of $5.49 is relatively consistent and does not appear to be influenced by any temporary sale-like conditions.
-Conclusion on Price Manipulation:
-
-Based on the price data, Voila may be engaging in price manipulation practices, where the price of the apples is first elevated to $6.49 and then temporarily reduced to $5.49 during the sale period, creating an illusion of a larger discount.
-This could be viewed as an artificial price increase before a sale, which aligns with the research question regarding whether prices are jacked up before a sale and lowered back to normal.
-Loblaws, by contrast, maintains a more stable pricing structure and does not appear to engage in such pricing practices.
-Overall, this analysis suggests that Voila may be artificially inflating prices ahead of sales to make the sale price appear more significant, while Loblaws maintains a consistent price strategy without such fluctuations.
-Further statistical analysis, such as testing for significant price changes before and after sales, would be necessary to substantiate this hypothesis more robustly.
+	1.	Price Comparison Between Vendors:
+	•	NoFrills offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items.
+This indicates a $1.00 decrease in price during the sale period, suggesting that NoFrills is promoting a sale by reducing the price from a regular price point.
+	•	Loblaws, on the other hand, has a consistent current price of $5.49 for all data points, which is generally lower than the sale price at NoFrills.
+Additionally, Loblaws’ old price fluctuates slightly around $5.49 to $6.58, but the regular price appears stable without the promotional fluctuations observed at NoFrills.
+	2.	Sale Behavior Analysis at NoFrills:
+	•	At NoFrills, the old price for the apples is consistently set at $6.49 before being reduced to $5.49 during sale periods.
+This consistent pricing behavior suggests that NoFrills applies a fixed discount strategy, where the regular price is consistently reduced by $1.00 during promotional events.
+	•	The stable difference between the sale and non-sale prices raises questions about whether NoFrills may be inflating prices just before the sale period to create the perception of a significant discount.
+The predictable pattern of a $1.00 reduction during sales could indicate a structured approach to pricing, potentially aimed at attracting price-sensitive customers through a visible discount.
+	3.	Price Consistency at Loblaws:
+	•	Loblaws displays a more stable pricing pattern, with the current price consistently at $5.49 and only slight variation in old prices.
+Unlike NoFrills, Loblaws does not engage in visible promotional adjustments, indicating a transparent and steady pricing strategy that does not rely on artificial price fluctuations.
+	•	The absence of sale-driven price changes suggests that Loblaws might prioritize a straightforward pricing model that could appeal to customers who value consistency over promotional sales.
+	4.	Conclusion on Price Manipulation:
+	•	The analysis indicates that NoFrills may engage in a pricing strategy where prices are elevated to $6.49 before sales and then reduced back to a sale price of $5.49, giving the impression of a more substantial discount.
+This strategy could be perceived as artificial price manipulation, potentially intended to create a psychological effect on customers.
+	•	In contrast, Loblaws maintains a consistent price, which may be perceived as more transparent and consumer-friendly, as it avoids the potential illusion of savings through price fluctuations.
 
 # Discussion
-The analysis of price data for **Apples Ambrosia** from **Voila** and **Loblaws** provides insight into pricing strategies across two grocery vendors.
-Specifically, the research question—whether prices are artificially increased just before a sale, only to be lowered back down—was examined by comparing **current prices** and **old prices** during sale periods at Voila and regular pricing at Loblaws.
 
-## Price Manipulation at Voila
+## Price Manipulation at NoFrills
 
-The results suggest that **Voila** may be engaging in a pricing strategy that aligns with the hypothesis of artificially inflating prices ahead of a sale.
-At Voila, the **old price** for the apples is consistently **$6.49**, which is advertised as the regular price.
-However, the **current price** during the sale period is consistently lower at **$5.49**, resulting in a fixed **$1.00 discount** across all sale entries.
-This consistent reduction implies that **Voila's sales** may be structured to create the illusion of a larger price cut, as prices are raised to **$6.49** just before the sale and then reduced to the normal price of **$5.49** during the sale.
+The findings suggest that NoFrills may engage in an artificial pricing strategy, where prices are raised before a sale period to create a perceived discount when they are subsequently lowered.
+This strategy involves advertising a stable $6.49 as the “regular price” before reducing it by $1.00 during sales.
+While this reduction represents a legitimate discount, the consistency of the sale pattern may lead consumers to believe they are receiving a more substantial benefit than they might in a transparent pricing model.
 
-This pricing behavior may lead customers to believe they are receiving a more significant discount than they actually are.
-Such price adjustments raise questions about transparency in pricing and the ethical implications of using these sales strategies.
-The strategy of artificially inflating prices before a sale is a tactic often criticized in retail, as it may mislead consumers into thinking they are getting better deals than they truly are.
-While the **$1.00 discount** in the sale period does represent a legitimate price reduction, it may not be as substantial as it appears when compared to the price manipulation in the lead-up to the sale.
+This approach could influence customer perceptions, potentially encouraging them to take advantage of “discounted” prices that may only appear significant due to prior price increases.
+The ethical implications of this tactic are worth noting, as customers may be misled about the actual value they are receiving, potentially impacting their trust in the brand.
 
@@
 
-In contrast, **Loblaws** maintains a more consistent pricing structure.
-The **current price** of **$5.49** for **Apples Ambrosia** is stable across all observations, and there is no recorded **sale** price or fluctuation in price over time.
-The **old price** at Loblaws varies slightly, but the difference between the **current** and **old price** is not as pronounced as at Voila.
-This suggests that Loblaws does not engage in the same type of artificial price elevation seen at Voila.
-Their pricing appears more transparent, without the pre-sale price hikes that would create the illusion of larger discounts.
-
-The stable pricing at **Loblaws** could be interpreted as a more consumer-friendly approach, with less likelihood of misleading customers about the "value" of sales.
-In fact, this could suggest that **Loblaws** may not use sales as a primary strategy to attract customers, but instead maintains competitive and consistent pricing for everyday purchases.
-While the price at Loblaws is lower than the sale price at Voila, it is important to consider that Loblaws may be applying a different pricing strategy based on its market positioning, competitive environment, and customer base.
+Loblaws appears to adopt a straightforward pricing strategy, with no apparent sale-driven adjustments.
+The stable $5.49 current price does not fluctuate, and there are no artificial discounts or price hikes before sales.
+This consistent approach might build customer trust, as customers can expect the same pricing without worrying about artificial discounts.
 
@@
 
-From a consumer perspective, the price manipulation strategy observed at Voila may have broader implications.
-Customers who are aware of the practice may become skeptical of sales and discounts, leading them to question the legitimacy of promotional offers.
-Such skepticism could erode trust in the vendor and result in a loss of customer loyalty.
-On the other hand, the consistent pricing at Loblaws could build consumer trust, as shoppers may feel confident that the price they see is the price they pay, without the manipulation of artificial sales tactics.
-
-However, it is also worth noting that the **$1.00 price difference** observed between **sale** and **regular price** is relatively modest and might not drastically affect consumer behavior unless similar strategies are applied to a wide range of products.
-Thus, while the price manipulation issue is noteworthy, it may not be the sole factor driving purchasing decisions in the broader context of grocery shopping, where convenience, availability, and other factors also play a significant role.
+The pricing strategy at NoFrills could lead some consumers to feel misled if they recognize the pattern of artificially inflated prices before sales.
+In contrast, the transparent pricing at Loblaws might foster consumer loyalty, as shoppers could perceive it as a more honest approach.
+The modest $1.00 difference between the “sale” and “regular” prices at NoFrills might not heavily impact purchasing behavior, but awareness of these strategies can empower consumers to make informed choices.
 
@@
 
-While this study provides useful insights, there are limitations to the analysis.
-The dataset only includes a limited range of data points for **Apples Ambrosia** across two vendors, which may not fully represent the broader pricing strategies employed by the vendors for all product categories.
-Further research would benefit from a more comprehensive dataset, encompassing a variety of products, seasons, and vendors.
-This would allow for a more robust understanding of pricing practices across different contexts.
+While this study provides useful insights, it is limited by the narrow scope of focusing on a single product across two vendors.
+Future studies could expand to include more products, seasonal data, and a wider range of vendors.
+Additionally, statistical testing of price differences could help confirm the significance of the observed patterns, providing stronger evidence regarding potential price manipulation.
 
-Moreover, additional statistical analysis could be conducted to test whether the observed price differences are statistically significant.
-This would help quantify the extent to which prices are artificially inflated before sales and provide more concrete evidence for or against the practice.
+# Conclusion
 
-## Conclusion
-
-In conclusion, the analysis indicates that **Voila** may be engaging in artificial price elevation ahead of sales, creating an illusion of a larger discount when the price is lowered back to normal during the sale period.
-**Loblaws**, in contrast, maintains a consistent and stable pricing strategy.
-While this raises important questions about transparency and fairness in pricing practices, it also provides an opportunity for consumers to make more informed decisions by understanding how sale prices may be manipulated.
-Further research would be needed to investigate this practice across a broader range of products and vendors, with an emphasis on consumer perceptions and market dynamics.
+In conclusion, the analysis suggests that NoFrills may utilize price elevation tactics before sales to create an illusion of larger discounts, while Loblaws maintains a stable and transparent pricing model.
+Understanding these pricing strategies allows consumers to make more informed purchasing decisions.
+Further research into broader pricing practices would enhance the insights from this analysis, potentially examining consumer responses to these strategies.
 

### 2024-11-15 paper/paper.qmd
Update paper.qmd

--- 2024-11-15 paper/paper.qmd
@@
 ## Overview
-
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-print(data)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-data_long <- data %>%
-  mutate(price_type = ifelse(!is.na(other), "Sale", "Regular")) %>%
-  pivot_longer(cols = c(current_price, old_price), values_to = "price")
-
-# Create the line plot
-ggplot(data_long, aes(x = product_vendor, y = price, color = price_type, linetype = price_type, group = interaction(price_type, product_vendor))) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(~product_vendor) + # Separate lines for each vendor
-  labs(title = "Price Comparison Over Time for Apples Ambrosia",
-       x = "Vendor", 
-       y = "Price", 
-       color = "Price Type", 
-       linetype = "Price Type") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
+We use the statistical programming language R [@citeR].
+The data used in this paper came from the [@hammer].
+Following @tellingstories, we considerto capture the pricing dynamics of Ambrosia apples across different vendors, we extracted relevant data from a larger dataset containing multiple products and vendors.
+The initial dataset included information from seven vendors, each offering various products.
+Our focus was on the specific product, "Ambrosia Apples," which was sold by two vendors—NoFrills and Loblaws.From this comprehensive dataset, we filtered out all other products, isolating only the entries related to Ambrosia apples.
+The extracted data included key information such as the product ID, product name, current price, and old price.
+Additionally, a field labeled "other" contained further notes on sale conditions or discounts.
+The dataset also recorded timestamps, allowing us to track pricing changes over time.
 
 ## Measurement
+
 To investigate the potential for artificial price increases prior to sales events, this study employs a data-driven measurement approach to track and analyze price fluctuations over time.
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-loblaws_data
+#| tbl-cap: "Ambrosia Apples Data of Loblaws"
+#| warning: false
+
+data |>
+  filter(product_vendor == "Loblaws") |>
+  select(product_name,product_vendor,current_price,old_price,nowtime) |>      # Select the first 3 columns
+  kable(
+    col.names = c("Product Name","Product Vendor","Product Price(current)","Product Price(old)","Date"),
+    booktabs = TRUE
+  )
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-nofrills_data
+#| tbl-cap: "Ambrosia Apples Data of NoFrills"
+#| tbl-width: 2
+#| tbl-height: 2
+#| warning: false
+
+data |>
+  filter(product_vendor == "NoFrills") |>
+  select(product_name,product_vendor,current_price,old_price,nowtime) |>    
+  kable(
+    col.names = c("Product Name","Product Vendor","Product Price (current)","Product Price (old)","Date"),
+    booktabs = TRUE
+  )
+```
+
+The transition from real-world phenomena (e.g., pricing decisions and promotional strategies) to dataset entries was facilitated by these fields.Current and old prices reflected the vendors' listed prices at the time of data collection.
+Notes in the "other" field provided contextual information about whether the price was marked as a sale or if there were discounts.
+The "nowtime" column captured the exact timing of data collection, enabling us to analyze price fluctuations and assess sale validity over time.
+
+By focusing on two vendors and this specific product, we created a clean, targeted dataset shows in @tbl-loblaws_data and @tbl-nofrills_data that accurately reflects the pricing behavior for Ambrosia apples at NoFrills and Loblaws, laying the foundation for further analysis.
+
+## Outcome variables
+
+```{r}
+#| label: fig-compare_price_vendor
+#| fig-cap: Current and Old Prices of Ambrosia Apples Over Time
+#| echo: false
+#| warning: false
+#| message: false
+
+clean_data$nowtime <- as.POSIXct(clean_data$nowtime, format="%Y-%m-%d %H:%M:%S")
+
+clean_data$product_vendor <- factor(clean_data$product_vendor, levels = c("NoFrills", "Loblaws"))
+
+# Create the line plot
+ggplot(clean_data, aes(x = nowtime)) +
+  geom_line(aes(y = current_price, color = product_vendor, linetype = "Current Price")) +
+  geom_line(aes(y = old_price, color = product_vendor, linetype = "Old Price")) +
+  labs(title = "Current and Old Prices of Ambrosia Apples Over Time",
+       x = "Date",
+       y = "Price ($)",
+       color = "Vendor",
+       linetype = "Price Type") +
+  theme_minimal() +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme(legend.position = "right")
+```
+We compared the apple prices of nofrills and loblaws at different times (including the comparison of old prices, where NoFrills does not include any old prices because this vendor does not have any discounts for this product and always maintains the original price), which can be seen in @fig-compare_price_vendor.
+
 # Results
 Price Analysis of "Apples Ambrosia" Across Vendors
-The dataset consists of price data for Apples Ambrosia from two grocery vendors: Voila and Loblaws.
-For each vendor, prices are observed during periods labeled as "SALE" (Voila) and regular pricing periods (Loblaws).
+The dataset consists of price data for Apples Ambrosia from two grocery vendors: NoFrills and Loblaws.
+For each vendor, prices are observed during periods labeled as "SALE" (Loblaws) and regular pricing periods (NoFrills).
 The analysis investigates the following:
 
-Price Comparison Between Vendors:
+## Price Comparison Between Vendors
+Loblaws offers a current price for the apples \$1.21, with an old price of \$1.45 for all sale items.
+This suggests a $0.24 decrease in price during the sale period, indicating that Loblaws is promoting a sale by reducing the price from the regular price.
+NoFrills, on the other hand, has a consistent current price of $1.2 for all data points.
+This is the same as the sale price at Loblaws.
+Additionally, NoFrills has no old price since they didn't do any discounts.
+However, the regular price at NoFrills appears to be stable without the promotional fluctuations observed at Loblaws.
 
-Voila offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items.
-This suggests a $1.00 decrease in price during the sale period, indicating that Voila is promoting a sale by reducing the price from the regular price.
-Loblaws, on the other hand, has a consistent current price of $5.49 for all data points.
-This is lower than the sale price at Voila.
-Additionally, Loblaws’ old price is slightly higher, averaging around $5.49 to $6.58.
-However, the regular price at Loblaws appears to be stable without the promotional fluctuations observed at Voila.
-Sale Behavior Analysis at Voila:
-
-At Voila, the old price for the apples is $6.49, which is advertised as the regular price.
-The current price during the sale is lower by exactly $1.00, consistently across multiple observations.
-This consistent pricing pattern across the dataset suggests that Voila uses a fixed price reduction strategy during sales, offering a price decrease of $1.00 off the original price during promotions.
-An important observation is that the old price at Voila (before the sale) is consistently $6.49, which may suggest that prices were artificially elevated just ahead of the sale, only to be reduced back down to the normal sale price during the promotional period.
-The stable difference between the sale and non-sale prices raises a question about the pricing practices at Voila.
-Price Consistency at Loblaws:
-
-At Loblaws, no specific sale event is recorded, and the prices are relatively stable, with current prices consistently at $5.49 for "Apples Ambrosia." The old price for Loblaws varies slightly, but the regular price does not undergo significant fluctuations like the sales at Voila.
-This stability in pricing indicates that Loblaws may not engage in price manipulation strategies for sale periods as seen at Voila.
-The current price of $5.49 is relatively consistent and does not appear to be influenced by any temporary sale-like conditions.
-Conclusion on Price Manipulation:
-
-Based on the price data, Voila may be engaging in price manipulation practices, where the price of the apples is first elevated to $6.49 and then temporarily reduced to $5.49 during the sale period, creating an illusion of a larger discount.
-This could be viewed as an artificial price increase before a sale, which aligns with the research question regarding whether prices are jacked up before a sale and lowered back to normal.
-Loblaws, by contrast, maintains a more stable pricing structure and does not appear to engage in such pricing practices.
-Overall, this analysis suggests that Voila may be artificially inflating prices ahead of sales to make the sale price appear more significant, while Loblaws maintains a consistent price strategy without such fluctuations.
-Further statistical analysis, such as testing for significant price changes before and after sales, would be necessary to substantiate this hypothesis more robustly.
+## Conclusion: Price Manipulation Practices?
+Despite Loblaws advertising a price drop from \$1.45 to \$1.21 as part of a "sale," the price at NoFrills is consistently set at \$1.21, without any promotional claims.
+This raises suspicions that Loblaws may be engaging in price manipulation practices, where the advertised sale price is simply a temporary reduction to match NoFrills' regular price, rather than offering a genuine discount.
+These findings highlight potential concerns about transparency in pricing and suggest that consumers may be misled into believing they are receiving a larger discount than is actually being offered.
+Loblaws' "sale" price may not reflect a true reduction in value, but rather a strategic pricing move to attract customers by promoting the illusion of a better deal.
+This could be seen as a form of pricing manipulation, where the apparent discount is used to create the perception of a more significant bargain.
+Consumers should be cautious when interpreting sales promotions and be aware of the actual price comparisons across different vendors to make informed purchasing decisions.
 
@@
 
-## Conclusion
+## Correlation vs. Causation
+In this analysis, we have observed a correlation between the sale price at Loblaws and the regular price at NoFrills, suggesting that both prices are set at $1.21 during the sale period.
+However, it is important to note that correlation does not imply causation.
+While the two prices may appear to be related, this analysis does not establish a causal relationship between Loblaws' price reduction and NoFrills' pricing strategy.
+For example, it is possible that Loblaws adjusted its price temporarily to match NoFrills' pricing rather than making a true price reduction.
+A more comprehensive causal analysis would require additional data points, such as customer behavior during sale periods, or experiments where prices are manipulated independently, to determine whether Loblaws' pricing strategy is indeed a response to NoFrills or if there are other factors influencing both vendors' pricing.
 
-In conclusion, the analysis indicates that **Voila** may be engaging in artificial price elevation ahead of sales, creating an illusion of a larger discount when the price is lowered back to normal during the sale period.
-**Loblaws**, in contrast, maintains a consistent and stable pricing strategy.
-While this raises important questions about transparency and fairness in pricing practices, it also provides an opportunity for consumers to make more informed decisions by understanding how sale prices may be manipulated.
-Further research would be needed to investigate this practice across a broader range of products and vendors, with an emphasis on consumer perceptions and market dynamics.
+## Missing Data
+Our dataset consists of prices from two vendors (NoFrills and Loblaws), but the full dataset includes many more products and vendors.
+Missing data could be a potential concern in our analysis, especially if key information, such as prices for other vendors, was not available or if data points for certain time periods were omitted.
+Missing data can bias the analysis if it is not random; for example, if sales data is missing during peak shopping periods when prices are likely to be highest, this could result in a lower estimate of the actual price difference.
+In our case, the absence of historical prices for NoFrills or detailed pricing data for other vendors might limit the conclusions we can draw about the broader pricing landscape.
+To address this, we could perform sensitivity analyses to evaluate how the results change with different assumptions about the missing data, or use imputation techniques to fill in gaps where appropriate.
+
+## Sources of Bias
+There are several potential sources of bias in this study that could influence our results.
+First, selection bias could occur if the data points from Loblaws and NoFrills are not representative of all pricing periods, or if certain time frames (e.g., weekends or holidays) are overrepresented.
+Since our data focuses on a subset of vendors and products, it may not fully capture the diversity of pricing practices across all 7 vendors in the full dataset.
+Additionally, vendor-specific biases may also exist if either Loblaws or NoFrills have particular promotional strategies that differ from those of other vendors.
+For instance, Loblaws' advertised sales may be selectively applied to high-demand products like Ambrosia apples, while NoFrills might use a more consistent pricing approach across its entire range of produce.
+Furthermore, potential reporting biases could affect the accuracy of the data if certain pricing changes were not accurately recorded or if vendor promotions were not consistently reflected in the dataset.
+To mitigate these biases, future analyses could involve a broader set of vendors and products, as well as methods to ensure that the dataset is representative and balanced.
 

### 2024-11-15 paper/paper.qmd

--- 2024-11-15 paper/paper.qmd
@@
 ## Overview
-We use the statistical programming language R [@citeR]....
+We use the statistical programming language R [@citeR].
 The data used in this paper came from the [@hammer].
@@
 
-	1.	Price Comparison Between Vendors:
-	•	NoFrills offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items.
-This indicates a $1.00 decrease in price during the sale period, suggesting that NoFrills is promoting a sale by reducing the price from a regular price point.
-	•	Loblaws, on the other hand, has a consistent current price of $5.49 for all data points, which is generally lower than the sale price at NoFrills.
-Additionally, Loblaws’ old price fluctuates slightly around $5.49 to $6.58, but the regular price appears stable without the promotional fluctuations observed at NoFrills.
-	2.	Sale Behavior Analysis at NoFrills:
-	•	At NoFrills, the old price for the apples is consistently set at $6.49 before being reduced to $5.49 during sale periods.
-This consistent pricing behavior suggests that NoFrills applies a fixed discount strategy, where the regular price is consistently reduced by $1.00 during promotional events.
-	•	The stable difference between the sale and non-sale prices raises questions about whether NoFrills may be inflating prices just before the sale period to create the perception of a significant discount.
-The predictable pattern of a $1.00 reduction during sales could indicate a structured approach to pricing, potentially aimed at attracting price-sensitive customers through a visible discount.
-	3.	Price Consistency at Loblaws:
-	•	Loblaws displays a more stable pricing pattern, with the current price consistently at $5.49 and only slight variation in old prices.
-Unlike NoFrills, Loblaws does not engage in visible promotional adjustments, indicating a transparent and steady pricing strategy that does not rely on artificial price fluctuations.
-	•	The absence of sale-driven price changes suggests that Loblaws might prioritize a straightforward pricing model that could appeal to customers who value consistency over promotional sales.
-	4.	Conclusion on Price Manipulation:
-	•	The analysis indicates that NoFrills may engage in a pricing strategy where prices are elevated to $6.49 before sales and then reduced back to a sale price of $5.49, giving the impression of a more substantial discount.
-This strategy could be perceived as artificial price manipulation, potentially intended to create a psychological effect on customers.
-	•	In contrast, Loblaws maintains a consistent price, which may be perceived as more transparent and consumer-friendly, as it avoids the potential illusion of savings through price fluctuations.
+## Price Comparison Between Vendors
+Loblaws offers a current price for the apples \$1.21, with an old price of \$1.45 for all sale items.
+This suggests a $0.24 decrease in price during the sale period, indicating that Loblaws is promoting a sale by reducing the price from the regular price.
+NoFrills, on the other hand, has a consistent current price of $1.2 for all data points.
+This is the same as the sale price at Loblaws.
+Additionally, NoFrills has no old price since they didn't do any discounts.
+However, the regular price at NoFrills appears to be stable without the promotional fluctuations observed at Loblaws.
+
+## Conclusion: Price Manipulation Practices?
+Despite Loblaws advertising a price drop from \$1.45 to \$1.21 as part of a "sale," the price at NoFrills is consistently set at \$1.21, without any promotional claims.
+This raises suspicions that Loblaws may be engaging in price manipulation practices, where the advertised sale price is simply a temporary reduction to match NoFrills' regular price, rather than offering a genuine discount.
+These findings highlight potential concerns about transparency in pricing and suggest that consumers may be misled into believing they are receiving a larger discount than is actually being offered.
+Loblaws' "sale" price may not reflect a true reduction in value, but rather a strategic pricing move to attract customers by promoting the illusion of a better deal.
+This could be seen as a form of pricing manipulation, where the apparent discount is used to create the perception of a more significant bargain.
+Consumers should be cautious when interpreting sales promotions and be aware of the actual price comparisons across different vendors to make informed purchasing decisions.
 
@@
 
-# Conclusion
+## Correlation vs. Causation
+In this analysis, we have observed a correlation between the sale price at Loblaws and the regular price at NoFrills, suggesting that both prices are set at $1.21 during the sale period.
+However, it is important to note that correlation does not imply causation.
+While the two prices may appear to be related, this analysis does not establish a causal relationship between Loblaws' price reduction and NoFrills' pricing strategy.
+For example, it is possible that Loblaws adjusted its price temporarily to match NoFrills' pricing rather than making a true price reduction.
+A more comprehensive causal analysis would require additional data points, such as customer behavior during sale periods, or experiments where prices are manipulated independently, to determine whether Loblaws' pricing strategy is indeed a response to NoFrills or if there are other factors influencing both vendors' pricing.
 
-In conclusion, the analysis suggests that NoFrills may utilize price elevation tactics before sales to create an illusion of larger discounts, while Loblaws maintains a stable and transparent pricing model.
-Understanding these pricing strategies allows consumers to make more informed purchasing decisions.
-Further research into broader pricing practices would enhance the insights from this analysis, potentially examining consumer responses to these strategies.
+## Missing Data
+Our dataset consists of prices from two vendors (NoFrills and Loblaws), but the full dataset includes many more products and vendors.
+Missing data could be a potential concern in our analysis, especially if key information, such as prices for other vendors, was not available or if data points for certain time periods were omitted.
+Missing data can bias the analysis if it is not random; for example, if sales data is missing during peak shopping periods when prices are likely to be highest, this could result in a lower estimate of the actual price difference.
+In our case, the absence of historical prices for NoFrills or detailed pricing data for other vendors might limit the conclusions we can draw about the broader pricing landscape.
+To address this, we could perform sensitivity analyses to evaluate how the results change with different assumptions about the missing data, or use imputation techniques to fill in gaps where appropriate.
+
+## Sources of Bias
+There are several potential sources of bias in this study that could influence our results.
+First, selection bias could occur if the data points from Loblaws and NoFrills are not representative of all pricing periods, or if certain time frames (e.g., weekends or holidays) are overrepresented.
+Since our data focuses on a subset of vendors and products, it may not fully capture the diversity of pricing practices across all 7 vendors in the full dataset.
+Additionally, vendor-specific biases may also exist if either Loblaws or NoFrills have particular promotional strategies that differ from those of other vendors.
+For instance, Loblaws' advertised sales may be selectively applied to high-demand products like Ambrosia apples, while NoFrills might use a more consistent pricing approach across its entire range of produce.
+Furthermore, potential reporting biases could affect the accuracy of the data if certain pricing changes were not accurately recorded or if vendor promotions were not consistently reflected in the dataset.
+To mitigate these biases, future analyses could involve a broader set of vendors and products, as well as methods to ensure that the dataset is representative and balanced.
 

### 2024-11-15 paper/paper.qmd

@@
   filter(product_vendor == "Loblaws") |>
+  head(5) |>
   select(product_name,product_vendor,current_price,old_price,nowtime) |>      # Select the first 3 columns
@@
   filter(product_vendor == "NoFrills") |>
+  head(5) |>
   select(product_name,product_vendor,current_price,old_price,nowtime) |>    

### 2024-11-15 paper/paper.qmd
Update paper.qmd

@@
 ---
-title: "Sale difference between Loblaws and Voila"
+title: "Sale difference between Loblaws and NoFrills"
+sub-title: "Is the Sale Price Really a Discount? Analyzing Potential Pre-Sale Price Manipulation"
 author: 
@@
   - Yihang Xu
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/vietng04/sale_difference_Loblaws_and_Voila](https://github.com/vietng04/sale_difference_Loblaws_and_Voila)."
 date: today

### 2024-11-15 paper/paper.qmd
Update the paper

@@
 library(dplyr)
-library(tidyr)
-library(tibble)
 library(knitr)
@@
 ## Overview
-We use the statistical programming language R [@citeR].
+We use the statistical programming language R [@citeR].Libraries `tidyverse` [@citetidyverse], `knitr` [@citeknitr], `readr`[@citereadr] and `dplyr` [@citedplyr] were used for simulating, cleaning and testing.
+Graphics were made using `ggplot2` [@citeGG].
  The data used in this paper came from the [@hammer].

### 2024-11-15 scripts/00-simulate_data.R
Update the scripts

@@ -1,0 +1,49 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of Apple between two Vendors
+# Author: Doran Wang
+# Date: 14 Nov. 2024
+# Contact: doran.wang@mail.utoronto.ca
+# License: None
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `difference_sale` rproj
+
+#### Workspace setup ####
+library(tidyverse)
+set.seed(304)
+
+#### Simulate data ####
+# Vendors
+vendor <- c("NoFrills", "Loblaws")
+
+# Prices
+prices <- c(1.21, 1.54)
+
+# Create a dataset by randomly assigning vendors and prices, with conditions for Loblaws and NoFrills
+analysis_data <- tibble(
+  product = paste("Product", 1:50),  # Add "Product" to make it a character
+  vendor = sample(
+    vendor,
+    size = 50,
+    replace = TRUE,
+    prob = c(0.5, 0.5) # Rough vendor distribution
+  ),
+  current_price = sample(
+    prices,
+    size = 50,
+    replace = TRUE
+  ),
+  old_price = if_else(
+    vendor == "Loblaws" & current_price == 1.21, 
+    1.54,  # Assign 1.54 if current price is 1.21
+    NA_real_  # Otherwise, set old_price to NA
+  )
+)
+
+# For NoFrills, we don't need old_price, so we will set it as NA
+analysis_data <- analysis_data %>%
+  mutate(
+    old_price = if_else(vendor == "NoFrills", NA_real_, old_price)
+  )
+
+#### Save data ####
+write_csv(analysis_data, "data/simulated_data.csv")

### 2024-11-15 scripts/01-test_simulated_data.R

@@ -1,0 +1,74 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated apple price dataset.
+# Author: Doran Wang
+# Date: 14 Nov. 2024
+# Contact: doran.wang@mail.utoronto.ca
+# License: None
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `differece_sale` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+data <- read_csv("data/simulated_data.csv")
+
+# Test if the data was successfully loaded
+if (exists("data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+# Check if the dataset has 50 rows
+if (nrow(data) == 50) {
+  message("Test Passed: The dataset has 50 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 50 rows.")
+}
+
+# Check if the dataset has 4 columns
+if (ncol(data) == 4) {
+  message("Test Passed: The dataset has 4 columns.")
+} else {
+  stop("Test Failed: The dataset does not have  columns.")
+}
+
+# Check if all values in the 'product' column are unique
+if (n_distinct(data$product) == nrow(data)) {
+  message("Test Passed: All values in 'product' are unique.")
+} else {
+  stop("Test Failed: The 'product' column contains duplicate values.")
+}
+
+# Check if the 'vendor' column contains only valid vendor name
+valid_vendor <- c("NoFrills","Loblaws")
+
+if (all(data$vendor %in% valid_vendor)) {
+  message("Test Passed: The 'vendor' column contains only valid vendor names.")
+} else {
+  stop("Test Failed: The 'vendor' column contains invalid vendor names.")
+}
+
+# Check if the 'current_price' column contains only valid price
+valid_prices <- c(1.21,1.54,NA)
+
+if (all(data$current_price %in% valid_prices)) {
+  message("Test Passed: The 'current_price' column contains only valid prices.")
+} else {
+  stop("Test Failed: The 'current_price' column contains invalid prices.")
+}
+
+# Check if the 'old_price' column contains only valid price
+valid_prices <- c(1.54,NA)
+
+if (all(data$old_price %in% valid_prices)) {
[…]

### 2024-11-15 paper/paper.qmd
Update paper.qmd

@@
   - Yihang Xu
-thanks: "Code and data are available at: [https://github.com/vietng04/sale_difference_Loblaws_and_Voila](https://github.com/vietng04/sale_difference_Loblaws_and_Voila)."
+thanks: "Code and data are available at: [https://github.com/vietng04/sale-difference-Loblaws-and-NoFrills](https://github.com/vietng04/sale-difference-Loblaws-and-NoFrills)."
 date: today

### 2024-11-15 scripts/03-clean_data.R
delte clean_data.R

@@ -1,26 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw data from https://jacobfilipp.com/hammer/
-# Author: Doran Wang
-# Date: 14 Nov. 2024
-# Contact: doran.wang@mail.utoronto.ca
-# License: None
-# Pre-requisites: download the dataset hammer-2-processed.sqlite
-# Any other information needed? using the code in DB Browser for SQLite
-
-CREATE TABLE clean_data AS 
-SELECT 
-    p.id AS product_id,
-    p.vendor AS product_vendor, 
-	p.product_name AS product_name,
-	r.other AS other,
-    r.current_price AS current_price,
-	r.old_price AS old_price,
-	r.nowtime AS nowtime
-FROM 
-    product p
-JOIN 
-    raw r 
-    ON p.id = r.product_id
-WHERE 
-    r.current_price IS NOT NULL
-    AND r.product_id IN (67188, 31508);
